Analysis of transit delays in the Toronto Transit Commission: 2024

### 2024-09-27 paper/paper.qmd
Update paper.qmd

---
title: "Analysis of transit delays in the Toronto Transit Commission: 2024"
subtitle: "Insights from Jan 1 to Aug 31"
author: "Dong Jun Yoon"
thanks: "Code and data are available at: https://github.com/DJY1231/TTC-report-2024.git"
date: today
date-format: long
abstract: "This paper analyzes TTC delay data to understand patterns and causes of delays. It finds buses experience the most delays, especially on certain routes, while subways have frequent undocumented delays. Mechanical issues, operations, and diversions are most common delay causes. The analysis indicates the need to reduce bus delays through more vehicles or tactics like bus bridging on certain routes."
format: pdf
number-sections: true
bibliography: references.bib

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```
# Introduction
Having operated for over 100 years, the Toronto Transit Commission (TTC) has become one of the most visible and representative public transportation organizations in North America. It is well-known as the primary mode of public transportation in Toronto's daily life. Within the TTC, the bus network typically experiences higher daily ridership compared to streetcars, while subways accommodate the largest volume of passengers due to their central role in covering longer routes. Frequent riders often notice delays that arise on a daily basis. As delays become worse and services becoming increasingly inconsistent, it is important to examine the potential causes, allowing the TTC to identify and implement effective solutions.

@@

# Overview of datasets
```{r}
install.packages("kableExtra")

library(kableExtra)
library(dplyr)
library(knitr)

@@

```
Table 1: Sample of bus delay statistics; the first 10 rows of TTC bus delay over time: Jan 1 to Aug 31, 2024.

```{r}
library(kableExtra)
library(dplyr)
library(knitr)

@@

```{r}
library(kableExtra)
library(dplyr)
library(knitr)

@@

# Data Visualization {#sec-results}
```{r}
# Load the required libraries
library(ggplot2)
library(dplyr)

@@

```{r}
library(ggplot2)
library(dplyr)

@@

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

@@

One key insight from the paper is that buses tend to experience delays more frequently than subways or streetcars, occurring approximately 2.35 times more often than subway delays and 4.07 times more than streetcar delays. This is a significant finding, as it suggests that the bus system is less reliable compared to other modes of transportation, likely due to factors such as traffic conditions, mechanical issues, and operational difficulties.Another observation is the difference in the reasons for delays among the different transportation types. For buses and streetcars, mechanical and operational issues are the leading causes of delays, while for subways, the main causes are patron-related incidents and mechanical failures. This differentiation highlights how the nature of the transportation system influences the kinds of challenges they face, emphasizing the need for tailored solutions to address these issues.

>>>>

---
title: "Analysis of transit delays in the Toronto Transit Commission: 2024"
subtitle: "Insights from Jan 1 to Aug 31"
author: "Dong Jun Yoon"
thanks: "Code and data are available at: https://github.com/DJY1231/TTC-report-2024.git"
date: today
date-format: long
abstract: "This paper provides analyzes of TTC delay data to understand the causes of delays. It illustrates uncover patterns in transit delays, comparing delay causes, trends over time, and variations across different vehicle types. Buses tend to experience delays more frequently than subways or streetcars, occurring approximately 2.35 times more often than subway delays and 4.07 times more than streetcar delays. It suggests that the bus system is less reliable compared to other modes of transportation, likely due to factors such as traffic conditions, mechanical issues, and operational difficulties."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction
Having operated for over 100 years, the Toronto Transit Commission (TTC) has become one of the most visible and representative public transportation organizations in North America. It is well-known as the primary mode of public transportation in Toronto's daily life. Within the TTC, the bus network typically experiences higher daily ridership compared to streetcars, while subways accommodate the largest volume of passengers due to their central role in covering longer routes. Frequent riders often notice delays that arise on a daily basis. As delays become worse and services becoming increasingly inconsistent, it is important to examine the potential causes, allowing the TTC to identify and implement effective solutions.

@@

# Overview of datasets
```{r, echo=FALSE}
suppressMessages({
  install.packages("kableExtra")

  library(kableExtra)
  library(dplyr)
  library(knitr)
})

@@

kable(subset_data, format = "pipe", align = 'l') %>%
  kable_styling(full_width = F, position = "center")
```
Table 1: Sample of bus delay statistics; the first 10 rows of TTC bus delay over time: Jan 1 to Aug 31, 2024.

```{r, echo=FALSE}
library(kableExtra)
library(dplyr)
library(knitr)

@@

```{r, echo=FALSE}
library(kableExtra)
library(dplyr)
library(knitr)

@@

# Data Visualization {#sec-results}
```{r, echo=FALSE}
# Load the required libraries
library(ggplot2)
library(dplyr)

@@

```{r, echo=FALSE}
library(ggplot2)
library(dplyr)

@@

```{r, echo=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)

@@

One key insight from the paper is that buses tend to experience delays more frequently than subways or streetcars, occurring approximately 2.35 times more often than subway delays and 4.07 times more than streetcar delays. This is a significant finding, as it suggests that the bus system is less reliable compared to other modes of transportation, likely due to factors such as traffic conditions, mechanical issues, and operational difficulties. Another observation is the difference in the reasons for delays among the different transportation types. For buses and streetcars, mechanical and operational issues are the leading causes of delays, while for subways, the main causes are patron-related incidents and mechanical failures. This differentiation highlights how the nature of the transportation system influences the kinds of challenges they face, emphasizing the need for tailored solutions to address these issues.
